---
title: "Simple Linear Regression - R"
output: html_notebook
---

## Building Linear regression with Toyota data

Clear workspace and set work dir

```{r}
rm(list=ls(all=TRUE))
setwd("C:\\Users\\samanoh.FAREAST\\besantr\\linear-regression-lab")
```

Read data
```{r}
data = read.csv("Simple_Linear_Reg.csv")
str(data)
```

Remove ID column
```{r}
data$Id = NULL
str(data)
```

Get independent and target variables and build model
```{r}
x = data$Price
y = data$Age_06_15

lin_model = lm(formula = y~x, data = data)
lin_model
```
Plot the graph
```{r}
plot(Age_06_15 ~ Price, data = data)
abline(lm(Age_06_15 ~ Price, data=data)) # use '+' operator to combime line in R script
```

Get summary of Linear model
```{r}
summary(lin_model)
```

Predict using linear model
```{r}
my_prediction = predict(object = lin_model)
comparision = cbind(my_prediction, data$Age_06_15)
View(comparision)
```

## Calculate Metrics - RMSE

Install libraries
```{r}
#install.packages("MLmetrics")
library("MLmetrics")
```

calculate RMSE
```{r}
RMSE(y_pred = my_prediction, y_true = data$Age_06_15)
```

Perform outlier anlysis and repeat the process to get better metrics

```{r}
boxplot(data$Price)
```

```{r}
data$Price = ifelse(data$Price >= 22000, 22000, data$Price)
boxplot(data$Price)
```

Build linear model from outlied data
```{r}
new_model = lm(data$Age_06_15 ~ data$Price, data = data)
new_model
```
Predict using new model

```{r}
my_new_prediction = predict(object = new_model)
new_comparision = cbind(my_new_prediction, data$Age_06_15)
View(new_comparision)
```
Compare and plot new prediction
```{r}
plot(Age_06_15 ~ Price, data = data)
abline(new_model)
```
Get summary of new model
```{r}
summary(new_model)
```


