---
title: "Clusterning Sample"
author: "Mugilan"
date: "27 April 2018"
output: html_document
---
Making sure that all required all libraries are installed. I'm calling all libraries
```{r}
#install.packages("cluster")
library("cluster")

#install.packages("fpc")
library(fpc)

```
If you get error for installing packages, give direct path in the code.
"install.packages("fpc",repos = c("http://rstudio.org/_packages", "http://cran.rstudio.com"))"


Taking Iris as my data frame and removing the 5th column, since it has been 
already classified
```{r}
data<-iris[,-5]
str(data)
```

Looking for NA values
```{r}
sum(is.na(data))
```


Appltying Kmeans model to the data & plotting it
```{r}
cluster<-kmeans(x = data,centers = 3)

plotcluster(data,cluster$cluster)
```

```{r}
clusplot(data, cluster$cluster, color=TRUE, shade=TRUE,labels=2, lines=0)

```

```{r}
plot(data[1:2],cluster$cluster)
```

```{r}
y=iris$Species
table(y,cluster$cluster)

```


```{r}
plot(data[c("Sepal.Length", "Sepal.Width")], col=cluster$cluster)
```
Hierarchical clustering
```{r}
plot( hclust(dist(USArrests), "ave") )
```
