---
title: "Clustering - R"
output: html_notebook
---

# Perform clustering - K-means, k-means++, heirarchical on iris data set

Prepare the environment
```{r}
rm(list=ls(all=TRUE))
setwd("C:\\Users\\samanoh\\training\\besantr\\clustering-lab")
```

Install libraries
```{r}
#install.packages("cluster")
library("cluster")

#install.packages("fpc")
library("fpc")
```

Take iris dataset except last column (excluding target variable)
```{r}
# use "-" to drop columns
# iris[, -c(3:5)] - drop cols 3 to 5
# iris[,-c(3,5)] - drop cols 3 and 5
data <- iris[,-5]
str(data)
```

Looking for NA values
```{r}
sum(is.na(data))
```

Applying k-means model and plotting it
```{r}
set.seed(10)
cluster <- kmeans(x=data, centers = 3)
plotcluster(data,cluster$cluster)
```

Plot cluster - using principal component variation
```{r}
clusplot(data, cluster$cluster, color=TRUE, shade=TRUE,labels=2, lines=0)
```

Try to plot only 2 columns to understand distribution
```{r}
plot(data[1:2],cluster$cluster)
```

Plot confusion matrix to compare with target variables
```{r}
y=iris$Species
table(y,cluster$cluster)
```

```{r}
plot(data[c("Sepal.Length", "Sepal.Width")], col=cluster$cluster)
```

# K-means++ algorithm

Install libraries
```{r}
#install.packages("LICORS")
library("LICORS")
```

Build k-means++ model
```{r}
set.seed(10)
kmpp <- kmeanspp(data=iris[,-5], k=3, start="random", nstart = 10)
plotcluster(iris[,-5], kmpp$cluster)
```

Plot using Princil component variation - PCV
```{r}
clusplot(iris[,-5], kmpp$cluster, color=TRUE, shade=TRUE,labels=2, lines=0)
```

Plot confusion matrix
```{r}
y=iris$Species
table(y,kmpp$cluster)
```

# heirarchical clustering

Heirarchial clustering using US crime rate data of all states

```{r}
crimedata = USArrests
crimedata
```

```{r}
distancematrix = dist(crimedata)
# method = ave is for agloromative
heirarchical_cluster = hclust(distancematrix, method = "ave")
plot(heirarchical_cluster)

```

