---
title: "Logistic Regression - R"
output: html_notebook
---

# Building Logistic regression with Bankruptcy data

```{r}
rm(list = ls(all=TRUE))
setwd("C:\\Users\\samanoh\\training\\besantr\\logistic-regression-lab")
```


Use Bankruptcy data

```{r}
bdata = read.csv("Bankruptcy_data.csv")
```

Install Libraries
```{r}
#install.packages("DMwR")
library(DMwR)
```

Perform imputation
```{r}
sum(is.na(bdata))
ci_data = centralImputation(bdata)
sum(is.na(ci_data))
```

Get structure of imputed data
```{r}
str(ci_data)
```

CHange the target column to categorical
```{r}
ci_data$target = as.factor(as.character(ci_data$target))
str(ci_data)
```

Get summary of target data to compute probable cutoff
```{r}
summary(ci_data$target)
cutoff = 1765/34788
cutoff
```
Perform Logistic regression
```{r}
log_bdata = glm(formula = target~., data=ci_data, family = binomial)
```

Get summary of model
```{r}
summary(log_bdata)
```

Predict the model
```{r}
prediction_bdata = predict(log_bdata, data = ci_data, type = "response")
# type = "response" ensures to get logistic regression type instead of linear output
```

Round the output ot 3 digits and view the output
```{r}
View(round(prediction_bdata,3))
```

Apply the cutoff and change numeric to categorical value
```{r}
log_output = ifelse(prediction_bdata>cutoff,1,0)
log_output<- as.factor(as.character(log_output))
summary(log_output)
```
# Metrics


```{r}
#install.packages("MLmetrics")
library(MLmetrics)
```

Calculate categorical metrics - Recall

```{r}
Recall(log_output, ci_data$target)
```

Calculate categorical metrics - precision
```{r}
Precision(log_output, ci_data$target)
```

Calculate categorical metrics - F1 score
```{r}
F1_Score(log_output, ci_data$target)
```

Plot confusion matrix
```{r}
ConfusionMatrix(log_output, ci_data$target)
```

# Apply different cutoff and try to calcualte metrics for comparision

```{r}
log_output = ifelse(prediction_bdata>0.05,1,0)
log_output<- as.factor(as.character(log_output))
summary(log_output)

print("Recall : ")
Recall(log_output, ci_data$target)

print("Precision : ")
Precision(log_output, ci_data$target)

print("F1 Score : ")
F1_Score(log_output, ci_data$target)

print("Confusion matrix")
ConfusionMatrix(log_output, ci_data$target)
```

Calculate ROCR for cutoff
```{r}
library(ROCR)

# Predicting on the train data
predicted <- predict(log_bdata,type="response")
prob <- prediction(predicted,ci_data$target)

# Getting the true positive rate and false negative rate
tprfpr <- performance(prob, "tpr", "fpr")

# Plotting the true positive rate and false negative rate based on the threshold value
plot(tprfpr,col=rainbow(10), colorize=T, print.cutoffs.at=seq(0,1,0.05))
```