---
title: "Feature Engineering - R"
output:
  html_document:
    df_print: paged
---

Clear workspace and set current work directory

```{r}
rm(list = ls(all=TRUE))
setwd("C:\\Users\\samanoh.FAREAST\\besantr\\feature-engineering-lab")
```

Import Toyota regression file
```{r}
data = read.csv("Simple_Linear_Reg.csv", header = TRUE)
summary(data)
```

Getting structure of data
```{r}
str(data)
max(data$Price)
min(data$Price)
```

## Transformation

Log transformed data
```{r}
log_data = log(data)
max(log_data$Price)
min(log_data$Price)
```


## Standardization

Install packages - vegan
```{r}
#install.packages("vegan")
library("vegan")
```

Standardization using Range
```{r}
range_std = decostand(data,method = "range")
max(range_std$Price)
min(range_std$Price)
```

Standardization using Standard Deviation
```{r}
sd_std = decostand(data, method = "standardize")
max(sd_std$Price)
min(sd_std$Price)
```

Declare two vectors height and weight and combine them to form a dataframe
Note: In case two vectors are not equal length, shortest vector values repeats
```{r}
height = c(145,163,172,183,191)
weight = c(56000,67500,73845,97345,100000)
data1 = cbind(height, weight)
```

Perform feature engineering on the data frame
```{r}
range_data1 = decostand(data1, method="range")
sd_data1 = decostand(data1, method = "standardize")
log_data1 = log(data1)
```

Calculate standard deviation of a vector
```{r}
sd(height)
```


