---
title: "Naive Bayes in R"
output: html_notebook
---

# Naive bayes implementation in R

Set up the environment

```{r}
rm(list=ls(all=TRUE))
setwd("C:\\Users\\samanoh\\training\\besantr\\naive-bayes-lab")
```

Read data
```{r}
Bank <- read.table("Logistic.txt", sep=";", header = TRUE)
str(Bank)
```

Recode the column values
```{r}
Bank$outcome <- ifelse(Bank$y == "yes", 1, 0)
Bank$outcome <- as.factor(as.character(Bank$outcome))
str(Bank$outcome)
Bank <- Bank[,-17]

```

Check for missing values
```{r}
sum(is.na(Bank))
```

Removing/changing numerical to categorical data
```{r}
Bank$age <- as.factor(as.character(Bank$age))
Bank$balance <- NULL
Bank$day <- as.factor(as.character(Bank$day))
Bank$duration <- NULL
Bank$campaign <-as.factor(as.character(Bank$campaign))
Bank$pdays <- NULL
Bank$previous <- as.factor( as.character(Bank$previous))
```

Looking at structure of final data and summary
```{r}
 str(Bank)
summary(Bank)
```

Install libraries
```{r}
#install.packages("naivebayes")
library("naivebayes")
```

Build model
```{r}
naive_model = naive_bayes(x=Bank[,-14], y=Bank$outcome)
```

Predicting on the same data
```{r}
my_predictions = predict(naive_model, Bank[,-14])
summary(my_predictions)
```

Comparing the accuracy
```{r}
comparision = cbind(Actual = Bank$outcome, Predicted = my_predictions)
View(comparision)
```

Build confusion matrix
```{r}
table(Actual = Bank$outcome, Predicted = my_predictions)
```
# Error metrics
Recall value
```{r}
library(MLmetrics)
recall_value = Recall(y_true = Bank$outcome, y_pred = my_predictions)
recall_value
```

Precision

```{r}
precision_value = Precision(y_true = Bank$outcome, y_pred = my_predictions)
precision_value
```

F1 score
```{r}
f1_score = F1_Score(y_true = Bank$outcome, y_pred = my_predictions)
f1_score
```

Specificity
```{r}
specificity_score = Specificity(y_true = Bank$outcome, y_pred = my_predictions)
specificity_score
```

