---
title: "Bankruptcy project in R"
output:
  html_document:
    df_print: paged
---

Clean workspace

```{r}
rm(list=ls(all=TRUE))
setwd("C:\\Users\\samanoh\\training\\besantr\\bankruptcy-project")
```

Get Data and view summary

```{r}
data = read.csv("Bankruptcy_data.csv", header = TRUE)
str(data)
```

COnvert target variable to categorical

```{r}
data$target = as.factor(as.character(data$target))
data$ID = NULL
str(data)
```

# Perform Random forest to find important attributes

Check for missing values

```{r}
sum(is.na(data))
```

Perform Central imputation if missing values found

```{r}
library(DMwR)
```

```{r}
ci_data = centralImputation(data = data)
sum(is.na(ci_data))
```

Random forest

```{r}
library(caret)
library(randomForest)
```

```{r}
model = randomForest(target ~ ., data=ci_data, keep.forest=TRUE, do.trace=TRUE, ntrees=50)
```

Find importance from the model

```{r}
model$importance
```

Extract important attributes in data frame

```{r}
rf_imp_attr = data.frame(model$importance)
rf_imp_attr = data.frame(row.names(rf_imp_attr), rf_imp_attr[,1])
colnames(rf_imp_attr) = c("Attributes", "Importance")
rf_imp_attr = rf_imp_attr[order(rf_imp_attr$Importance, decreasing = TRUE),]
rf_imp_attr
```

Plot important attributes in graph

```{r}
varImpPlot(model)
```

# Select top 6 based on the importance

```{r}
imp_attr = as.character(rf_imp_attr$Attributes[1:6])
imp_attr
```

Select important attributes from actual data and perform iterations

```{r}
imp_data = data[imp_attr]
imp_data_full <- imp_data
imp_data_full$target <- data$target
str(imp_data_full)
```

# Perform outlier analysis - Box plot

First Feature - original
Attr27

```{r}
boxplot(imp_data_full$Attr27)$stats
```

First feature - outlied
Attr27

```{r}
imp_data_full$Attr27 = ifelse(imp_data_full$Attr27 >= 70000, NA, imp_data_full$Attr27)
imp_data_full$Attr27 = ifelse(imp_data_full$Attr27 <= -20000, NA, imp_data_full$Attr27)
boxplot(imp_data_full$Attr27)$stats
```

Second Feature - original
Attr34

```{r}
boxplot(imp_data_full$Attr34)$stats
```

Second feature - outlied
Attr34

```{r}
imp_data_full$Attr34 = ifelse(imp_data_full$Attr34 >= 150, NA, imp_data_full$Attr34)
imp_data_full$Attr34 = ifelse(imp_data_full$Attr34 <= -25, NA, imp_data_full$Attr34)
boxplot(imp_data_full$Attr34)$stats
```

Third attribute - original
Attr46

```{r}
boxplot(imp_data_full$Attr46)$stats
```

Third attribute - outlied
Attr46

```{r}
imp_data_full$Attr46 = ifelse(imp_data_full$Attr46 >= 125, NA, imp_data_full$Attr46)
imp_data_full$Attr46 = ifelse(imp_data_full$Attr46 <= -50, NA, imp_data_full$Attr46)
boxplot(imp_data_full$Attr46)$stats
```

Fourth attribute - original
Attr58

```{r}
boxplot(imp_data_full$Attr58)$stats
```

Fourth attribte - Outlied
Attr58

```{r}
imp_data_full$Attr58 = ifelse(imp_data_full$Attr58 >= 5, NA, imp_data_full$Attr58)
imp_data_full$Attr58 = ifelse(imp_data_full$Attr58 <= 0, NA, imp_data_full$Attr58)
boxplot(imp_data_full$Attr58)$stats
```

Fifth Attribute - original
Attr56

```{r}
boxplot(imp_data_full$Attr56)$stats
```

Fifth Attribute - outlied
Attr56

```{r}
imp_data_full$Attr56 = ifelse(imp_data_full$Attr56 >= 1, NA, imp_data_full$Attr56)
imp_data_full$Attr56 = ifelse(imp_data_full$Attr56 <= -4, NA, imp_data_full$Attr56)
boxplot(imp_data_full$Attr56)$stats
```

Sixth Attribute - original
Attr24

```{r}
boxplot(imp_data_full$Attr24)$stats
```

Sixth attribute - outlied
Attr24

```{r}
imp_data_full$Attr24 = ifelse(imp_data_full$Attr24 >= 4.5, NA, imp_data_full$Attr24)
imp_data_full$Attr24 = ifelse(imp_data_full$Attr24 <= -5, NA, imp_data_full$Attr24)
boxplot(imp_data_full$Attr24)
```

# Missing value treatment - KNN imputation

Check missing values

```{r}
sum(is.na(imp_data_full))
```

Import libraries

```{r}
library(DMwR)
```


Find Columns with more than 3 NA values and ignore (or select columns with <=3 null values)
(KNN imputation can't be performed for rows having values less than 3)

```{r}
imp_data_no_na = imp_data_full[rowSums(is.na(imp_data_full)) <= 3,]
str(imp_data_no_na)
```

# Train test split

Split imputed data
```{r}
set.seed(88)
trainrows = createDataPartition(y=imp_data_no_na$target, p=0.8, list = FALSE)
train_split = imp_data_no_na[trainrows,]
val_split = imp_data_no_na[-trainrows,]
```

Perform KNN imputation on train data

```{r}
train_split_0 = train_split[train_split$target==0,]
train_split_1 = train_split[train_split$target==1,]

train_split_0_knn = knnImputation(data = train_split_0, k=10)
train_split_1_knn = knnImputation(data = train_split_1, k=10)

sum(is.na(train_split_0_knn))
sum(is.na(train_split_0_knn))

# use distribution of target value 1 for val_split imputation since recall is the right metrics
val_split_knn = knnImputation(data = val_split, k=10, distData = train_split_1_knn)
sum(is.na(val_split_knn))

# Combine imputed train data
train_split_knn = rbind(train_split_0_knn, train_split_1_knn)
```

# Check variables where standard deviation is 0

Ignore such values to avoid bias

```{r}
sd_train=apply(train_split_knn,2,sd)
sort(sd_train,decreasing = TRUE)

# no SD=0 values found
```



# Build Model - Logistic regression

Build model with glm for logistic regression

```{r}
log_reg_model = glm(formula = target~., data = train_split_knn, family = binomial)
```

Predict the model from val_data

```{r}
log_reg_prediction = predict(log_reg_model, newdata=val_split_knn[-7], type="response")
```

View predicted output

```{r}
View(round(log_reg_prediction,3))
```

# Logistic regression Model - apply cutoff based on distribution

Identify cutoff

```{r}
summary(train_split_knn$target)

cutoff = 1411/27804
cutoff
```

Apply cutoff on prediction

```{r}
log_reg_output = ifelse(log_reg_prediction >= cutoff, 1, 0)
log_reg_output = as.factor(as.character(log_reg_output))
summary(log_reg_output)
View(log_reg_output)
```

# Metrics - Logistic regression without ROC

Import library
```{r}
library("MLmetrics")
```

Plot Confusion matrix
```{r}
confusionMatrix(log_reg_output, val_split_knn$target)
```

Calculate categorical metrics - Recall

```{r}
Recall(log_reg_output, val_split_knn$target)
```

Calculate categorical metrics - precision

```{r}
Precision(log_reg_output, val_split_knn$target)
```

Calculate categorical metrics - F1 score

```{r}
F1_Score(log_reg_output, val_split_knn$target)
```

# Logistic regression Model - apply cutoff based on ROC

Import library
```{r}
library("ROCR")
```

Calculate ROC

```{r}
rocr_pred = prediction(log_reg_prediction, val_split_knn$target)
rocr_perf = performance(rocr_pred, "tpr", "fpr")
plot(rocr_perf, colorize = TRUE)
```

Get ROC cutoff value


```{r}
#install.packages("Epi")
library(Epi)
```

```{r}
rc <- ROC(form = target~., data = val_split_knn, plot="sp") 
```

```{r}
## optimal combination
opt <- which.max(rowSums(rc$res[, c("sens", "spec")]))
## optimal cut-off point 
roc_cutoff = rc$res$lr.eta[opt]
roc_cutoff
```

Plotting the ROC cutoff in graph

```{r}
ROC(form = target~., data = val_split_knn, plot = "ROC", MX = TRUE)
```


Apply ROC cutoff on prediction

```{r}
log_reg_output = ifelse(log_reg_prediction >= roc_cutoff, 1, 0)
log_reg_output = as.factor(as.character(log_reg_output))
summary(log_reg_output)
```




# Metrics - Logistic regression with ROC

Import library
```{r}
library("MLmetrics")
```

Plot Confusion matrix
```{r}
confusionMatrix(log_reg_output, val_split$target)
```

Calculate categorical metrics - Recall

```{r}
Recall(log_reg_output, val_split$target)
```

Calculate categorical metrics - precision

```{r}
Precision(log_reg_output, val_split$target)
```

Calculate categorical metrics - F1 score

```{r}
F1_Score(log_reg_output, val_split$target)
```




# Build Model - KNN

Install libraries

```{r}
#install.packages("class")
library("class")
```

Perform KNN classification

```{r}
knn_class_prediction = knn(train = train_split_knn[-7], test = val_split_knn[-7], cl = train_split_knn$target, k=7 )
```


# Metrics - KNN

Finding accuracy

```{r}
#install.packages("gmodels")
library("gmodels")
```

Using crosstable for confusion matrix

```{r}
CrossTable(x = val_split_knn$target, y = knn_class_prediction, prop.chisq = FALSE)
```

Calculate Accuracy

```{r}
(6910+47)/7303
```

Install MLmetrics

```{r}
library("MLmetrics")
```

Calculate categorical metrics - Recall

```{r}
Recall(y_true = val_split_knn$target, y_pred = knn_class_prediction)
```

Calculate categorical metrics - precision

```{r}
Precision(y_true = val_split_knn$target, y_pred = knn_class_prediction)
```

Calculate categorical metrics - F1 score

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = knn_class_prediction)
```


# Build Model - Random Forest

Build model

Load Random forest library

```{r}
library(caret)
library(randomForest)
```

```{r}
rf_model = randomForest(target~., data=train_split_knn, keep.forest=TRUE, do.trace=TRUE, ntrees=50)
```

Predict validation data using random forest

```{r}
rf_prediction = predict(rf_model, val_split_knn[-7], type="response", norm.votes=TRUE)
```


# Metrics - Random Forest

Confusion Matrix

```{r}
confusionMatrix(data = rf_prediction, reference = val_split_knn$target,positive = "1")
```

Recall - Rf

```{r}
Recall(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Precision - rf

```{r}
Precision(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Sensitivity - Rf

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Specificity - Rf

```{r}
Specificity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

F1 Score - Rf

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = rf_prediction)
```


# Random forest - Iteration 2

```{r}
rf_model = randomForest(target~., data=train_split_knn, keep.forest=TRUE, ntrees=500, cutoff=c(0.75,0.25))
```

Predict validation data using random forest

```{r}
rf_prediction = predict(rf_model, val_split_knn[-7], type="response", norm.votes=TRUE)
```


# Metrics - Random Forest - iteration 2

Confusion Matrix

```{r}
confusionMatrix(data = rf_prediction, reference = val_split_knn$target,positive = "1")
```

Recall - Rf

```{r}
Recall(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Precision - rf

```{r}
Precision(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Sensitivity - Rf

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Specificity - Rf

```{r}
Specificity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

F1 Score - Rf

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = rf_prediction)
```

# Random forest - Iteration 3

```{r}
#install.packages("ROSE")
library(ROSE)
```


```{r}

#Due to lot of imbalance in the class , balanced it using ovun.sample function
data_balanced_both <- ovun.sample(target ~ ., data = train_split_knn, method = "both", p=0.5, seed = 1)$data
table(data_balanced_both$target)
```

```{r}
rf_model = randomForest(target~., data=data_balanced_both, keep.forest=TRUE, ntrees=500)
```

Predict validation data using random forest

```{r}
rf_prediction = predict(rf_model, val_split_knn[-7], type="response", norm.votes=TRUE)
```

# Metrics - Random Forest - iteration 3

Confusion Matrix

```{r}
confusionMatrix(data = rf_prediction, reference = val_split_knn$target,positive = "1")
```

Recall - Rf

```{r}
Recall(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Precision - rf

```{r}
Precision(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Sensitivity - Rf

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

Specificity - Rf

```{r}
Specificity(y_true = val_split_knn$target, y_pred = rf_prediction)
```

F1 Score - Rf

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = rf_prediction)
```


# Build Model - CART

Install libraries

```{r}
#install.packages("rpart")
library(rpart)
```

Build CART model

```{r}
cart_model = rpart(train_split_knn$target~.,data = train_split_knn)
summary(cart_model) 
```


Predict using CART

```{r}
cart_prediction = predict(cart_model, val_split_knn[-7], type = "class")
```


# Metrics - CART

Confusion Matrix

```{r}
confusionMatrix(data = cart_prediction, reference = val_split_knn$target,positive = "1")
```


Calculate categorical metrics - Recall

```{r}
Recall(y_true = val_split_knn$target, y_pred = cart_prediction)
```


Calculate categorical metrics - precision

```{r}
Precision(y_true = val_split_knn$target, y_pred = cart_prediction)
```


Calculate categorical metrics - sensitivity

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = cart_prediction)
```


Calculate categorical metrics - specificity

```{r}
Specificity(y_true = val_split_knn$target, y_pred = cart_prediction)
```


Calculate categorical metrics - F1 Score

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = cart_prediction)
```




# Build Model - C5.0

Install packages

```{r}
#install.packages("C50")
library("C50")
```


Build model C5.0

```{r}
c50_model = C5.0(x = train_split_knn[-7], y = train_split_knn$target)
summary(c50_model)
```

Predict model on validation data 

```{r}
c50_prediction = predict(c50_model, newdata=val_split_knn[-7])
```




# Metrics - C5.0

Confusion Matrix

```{r}
confusionMatrix(data = c50_prediction, reference = val_split_knn$target, positive = "1")
```


Calculate categorical metrics - Recall

```{r}
Recall(y_true = val_split_knn$target, y_pred = c50_prediction)
```


Calculate categorical metrics - precision

```{r}
Precision(y_true = val_split_knn$target, y_pred = c50_prediction)
```


Calculate categorical metrics - sensitivity

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = c50_prediction)
```


Calculate categorical metrics - specificity

```{r}
Specificity(y_true = val_split_knn$target, y_pred = c50_prediction)
```


Calculate categorical metrics - F1 Score

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = c50_prediction)
```


# Build Model - Naive Bayes

Install Libraries

```{r}
#install.packages("naivebayes")
library("naivebayes")
```

Build model

```{r}
naive_model = naive_bayes(x = train_split_knn[-7], y = train_split_knn$target)
summary(naive_model)
```

Predict model

```{r}
naive_prediction = predict(naive_model, newdata = val_split_knn[-7])
summary(naive_prediction)
```


# Metrics - Naive Bayes


Confusion Matrix

```{r}
confusionMatrix(data = naive_prediction, reference = val_split_knn$target, positive = "1")
```


Calculate categorical metrics - Recall

```{r}
Recall(y_true = val_split_knn$target, y_pred = naive_prediction)
```


Calculate categorical metrics - precision

```{r}
Precision(y_true = val_split_knn$target, y_pred = naive_prediction)
```


Calculate categorical metrics - sensitivity

```{r}
Sensitivity(y_true = val_split_knn$target, y_pred = naive_prediction)
```


Calculate categorical metrics - specificity

```{r}
Specificity(y_true = val_split_knn$target, y_pred = naive_prediction)
```


Calculate categorical metrics - F1 Score

```{r}
F1_Score(y_true = val_split_knn$target, y_pred = naive_prediction)
```


